### 2014-12-14 Rebuttal.txt

Some of the questions raised by the first three reviewers are addressed in the
comments of the Reviewer D.

We have submitted the source code of our Coq development as a suplement to our
paper in the submission website. It can also be obtained from the public repository
http://github.com/vmurali/SeqConsistency.

In terms of code-reuse, we have structured the proofs modularly as described in
the paper. Thus, the proofs of the different components can be reused while
changing the other components. For example, the proof of the cache-coherence
protocol can be reused as is for verifying other (non-sequentially consistent)
systems. Moreover, the proof of the cache-coherence is itself modular and can
be adopted to other protocols (like MESI, MOSI); The proofs for Lemmas 2 to 5
will change, but the lemmas correspond to at most 3 components (and states) and
hence can be easily proved (even using techniques like model-checking).

We agree with Reviewer-D that liveness is an important non-trivial property
that has to be proven for a cache-coherence protocol. The example of
ParentSendReq is an example of a voluntary rule, and the policy to fire such a
rule does indeed affect the liveness. In our development, we haven't proved
full liveness (which includes the liveness of the processor), but we have
proved the following (stated informally):
a) Any request (or response) that a cache wants to send can eventually be sent
(either to its parent or to its children)
b) Under the assumption that ParentSendReq rule (sending downgrade request from
a parent to its children) will fire eventually, whenever the parent has an
upgrade request from a child, every request will eventually get a response.

Thus, the only extraneous condition we impose is that a parent should "decide"
to send a downgrade request to (some of) its children whenever it sees an
upgrade request from one of its children. Eventually, such a request will be
sent via ParentSendReq rule.

We believe such a modular technique for verifying full-system correctness with
respect to SC will be useful for hardware designers/verification-engineers to
independently verify different components of the system. We do not model non-SC
systems (like ARM/POWER), which exhibit behaviors described by Reviewer-C
(involving shared store buffers, committing non-verified load values, etc).
Reviewer-C, please note that the idea of verification loads does indeed ensure
correctness, even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. In fact,
our final theorem ensures that this is indeed the case, otherwise the
speculative system will not be an implementation of the SC reference system.

Reviewer-B please note that our SC model is exactly as specified by Lamport.
Also please note that by local buffers in Figure 2, we do not mean just store
buffers; it includes all local data accessible to the processor.

Thanks for notifying us of other related work (Hu et.al.'s "Decidable SC"
SPAA'03, Qadeer's "Verifying SC by model-checking" TPDS'03, Sorin's
"Verification-aware Architectures" PACT'07, "Lamport Clocks: Verifying
Directory Cache-Coherence Protocol" SPAA'98 and Moore's "ACL2 proof of cache
coherence" CAV'98), and for pointing the other bugs in our related work
section.

Reviewer-B's concerns:
- Ideal processor and memory: as described in that paragraph - memory responds
  instantaneously and processor executes atomically without speculation
- Bluespec served as the model: The motivation for LTS comes from using
  Bluespec to describe hardware, which is similar to our LTS formalization.
- Bluespec's competitive performance can be seen from [Dave et al. 2005], and
  other papers using Bluespec designs.
- Peek-beneath abstraction boundaries: To prove the full-system, the
  individual components are treated as black-boxes, completely captured by their
  interface specification.
- Good-ideas from worlds of PL semantics: Using LTSes as a precise hardware
  specification.
- Infinite families of hardware design ... resist reduction to finite-state
  models: reduction is applied to 'infinite families'
- Figure 2: logical view of system as 3-components: It is more-or-less accepted
  in the architecture community.
- Refining vs Equivalent: Throughout the paper, we are showing only refinement
  in one direction, though you are correct that in this particular example they
  are equivalent.
- Finer-grained architectural details: as mentioned in that paragraph (like
  register-renaming)
- Main-memory still exposes atomic loads and stores: Most architectures (including
  POWER and ARM) can be modeled using a memory that responds to loads and stores
  instantaneously, though the response (or request) can sit in a buffer between
  the memory and the processor, leading to relaxed memory models.

### 2014-12-15 Rebuttal.txt

Some of the questions raised by the first three reviewers are addressed in the
comments of the Reviewer D.

We have submitted the source code of our Coq development as a supplement to our
paper in the submission website (we didn't give any link to preserve
anonymity).

We have structured the proofs modularly as described in the paper. Thus, the
proofs of the different components can be reused while changing the other
components. For example, the proof of the cache-coherence protocol can be
reused as is for verifying other (non-sequentially consistent) systems.
Moreover, the proof of the cache-coherence is itself modular and can be adopted
to other protocols (like MESI, MOSI); The proofs for Lemmas 2 to 5 will change,
but the lemmas correspond to at most 3 components (and states) and hence can be
easily proved (even using techniques like model-checking).

We agree with Reviewer-D that proving liveness is important and non-trivial.
The example of ParentSendReq is an example of a voluntary rule, and the policy
to fire such a rule does indeed affect the liveness. In our development, we
haven't proved full liveness of the system (which includes the liveness of the
processor), but we have proved the following which specifies the liveness of
the cache-coherence protocol:
a) Any request (or response) that a cache wants to send can eventually be sent
(either to its parent or to its children)
b) Under the assumption that a parent will eventually "decide" to send a
downgrade request to its children (using ParentSendReq rule) whenever it sees
an upgrade request from a child (as opposed to ignoring such an upgrade request
from a child forever), every request sent to either the parent or to a child
will eventually get a response.

We chose "halting" for pedagogical reasons in the paper; in our actual Coq
development, we use the more traditional trace-equivalence definition requiring
inter-convertible states. It is worth noting that any safety property can be
reduced to the problem of halting.

We believe such a modular technique for verifying full-system correctness with
respect to SC will be useful for hardware designers/verification-engineers to
independently verify different components of the system. We can't be sure that
everything will generalize effortlessly to the trickiest relaxed memory models
(like those exhibited by ARM/POWER as pointed by Reviewer-C because of no
verification loads and/or shared store buffers), but it would have been hard to
figure out the basic framework if we started on more complex examples first.
We still think our example is quite complex and interesting, which Reviewer-D
seems to agree with.

Reviewer-C, please note that verification loads do indeed ensure correctness.
This is true even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. Our final
theorem ensures this; otherwise the speculative system will not be an
implementation of the SC reference system.

@@

Reviewer-B please note that our SC model is exactly as specified by Lamport.
Also please note that by local buffers in Figure 2, we do not mean just store
buffers; it includes all local data accessible to the processor.

Thanks for notifying us of other related work (Hu et.al.'s "Decidable SC"
SPAA'03, Qadeer's "Verifying SC by model-checking" TPDS'03, Sorin's
"Verification-aware Architectures" PACT'07, "Lamport Clocks: Verifying
Directory Cache-Coherence Protocol" SPAA'98 and Moore's "ACL2 proof of cache
coherence" CAV'98), and to Reviewer-B for pointing the other bugs in our
related work section.

Reviewer-B's concerns:
- Ideal processor and memory: as described in that paragraph - memory responds
  instantaneously and processor executes atomically without speculation
- Bluespec served as the model: The motivation for LTS comes from using
  Bluespec to describe hardware, which is similar to our LTS formalization.
- Bluespec's competitive performance can be seen from [Dave et al. 2005], and
  other papers using Bluespec designs.
- Peek-beneath abstraction boundaries: To prove the full-system, the
  individual components are treated as black-boxes, completely captured by their
  interface specification.
- Good-ideas from worlds of PL semantics: Using LTSes as a precise hardware
  specification.
- Infinite families of hardware design resist reduction to finite-state
  models: reduction is applied to 'infinite families'
- Figure 2: logical view of system as 3-components: It is more-or-less accepted
  in the architecture community.
- Refining vs Equivalent: Throughout the paper, we are showing only refinement
  in one direction, though you are correct that in this particular example they
  are equivalent.
- Finer-grained architectural details: as mentioned in that paragraph (like
  register-renaming)
- Main-memory still exposes atomic loads and stores: Most architectures (including
  POWER and ARM) can be modeled using a memory that responds to loads and stores
  instantaneously, though the response (or request) can sit in a buffer between
  the memory and the processor, leading to relaxed memory models.

### 2014-12-15 Rebuttal.txt

Reviewer-C, please note that verification loads do indeed ensure correctness.
This is true even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. Our final
theorem ensures this; otherwise the speculative system will not be an
implementation of the SC reference system.

Reviewer-B please note that our SC model is exactly as specified by Lamport.
Also please note that by local buffers in Figure 2, we do not mean just store
buffers; it includes all local data accessible to the processor.

Thanks for notifying us of other related work (Hu et.al.'s "Decidable SC"
SPAA'03, Qadeer's "Verifying SC by model-checking" TPDS'03, Sorin's
"Verification-aware Architectures" PACT'07, "Lamport Clocks: Verifying
Directory Cache-Coherence Protocol" SPAA'98 and Moore's "ACL2 proof of cache
coherence" CAV'98), and to Reviewer-B for pointing the other bugs in our
related work section.

Reviewer-B's concerns:
- Ideal processor and memory: As described in that paragraph - memory responds
  instantaneously and processor executes atomically without speculation
- Bluespec served as the model: The motivation for LTS comes from using
  Bluespec to describe hardware, which is similar to our LTS formalization.
- Bluespec's competitive performance can be seen from [Dave et al. 2005], and
  other papers using Bluespec designs.
- Peek-beneath abstraction boundaries: To prove the full-system, the
  individual components are treated as black-boxes, completely captured by their
  interface specification.
- Good-ideas from the world of PL semantics: Using LTSes as a precise hardware
  specification.
- Infinite families of hardware design resist reduction to finite-state
  models: reduction is applied to 'infinite families'
- Figure 2: logical view of system as 3-components: It is more-or-less accepted
  in the architecture community.
- Refining vs Equivalent: Throughout the paper, we are showing only refinement
  in one direction, though you are correct that in this particular example they
  are equivalent.
- Finer-grained architectural details: As mentioned in that paragraph (like
  register-renaming).
- Main-memory still exposes atomic loads and stores: Most architectures (including
  POWER and ARM) can be modeled using a memory that responds to loads and stores
  instantaneously, though the response (or request) can sit in a buffer between
  the memory and the processor, leading to relaxed memory models.

### 2014-12-15 Rebuttal.txt

We believe such a modular technique for verifying full-system correctness with
respect to SC will be useful for hardware designers/verification-engineers to
independently verify different components of the system. We can't be sure that
everything will generalize effortlessly to the trickiest relaxed memory models
(like those exhibited by ARM/POWER as pointed by Reviewer-C because of no
verification loads and/or shared store buffers), but it would have been hard to
figure out the basic framework if we started on more complex examples first.
We still think our example is quite complex and interesting, which Reviewer-D
seems to agree with.

Reviewer-C, please note that verification loads do indeed ensure correctness.
This is true even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. Our final
theorem ensures this; otherwise the speculative system will not be an
implementation of the SC reference system.

Reviewer-B please note that our SC model is exactly as specified by Lamport.
Also please note that by local buffers in Figure 2, we do not mean just store
buffers; it includes all local data accessible to the processor.

Thanks for notifying us of other related work (Hu et.al.'s "Decidable SC"
SPAA'03, Qadeer's "Verifying SC by model-checking" TPDS'03, Sorin's
"Verification-aware Architectures" PACT'07, "Lamport Clocks: Verifying
Directory Cache-Coherence Protocol" SPAA'98 and Moore's "ACL2 proof of cache
coherence" CAV'98), and for pointing the other bugs in our related work
section.

### 2014-12-15 Rebuttal.txt

Some of the questions raised by the reviewers are addressed by Reviewer-D.

@@

We have submitted the source code of our Coq development as a supplement to our
paper in the submission website (we didn't give any link to preserve
anonymity).

We have structured the proofs modularly as described in the paper. Thus, the
proofs of the different components can be reused while changing the other
components. For example, the proof of the cache-coherence protocol can be
reused as is for verifying other (non-sequentially consistent) systems.
Moreover, the proof of the cache-coherence is itself modular and can be adopted
to other protocols (like MESI, MOSI); The proofs for Lemmas 2 to 5 will change,
but the lemmas correspond to at most 3 components (and states) and hence can be
easily proved (even using techniques like model-checking).

We agree with Reviewer-D that proving liveness is important and non-trivial.
The example of ParentSendReq is an example of a voluntary rule, and the policy
to fire such a rule does indeed affect the liveness. In our development, we
haven't proved full-system liveness (which includes the liveness of the
processor), but we have proved the liveness of the cache-coherence protocol, in
particular, the following:
a) Any request (or response) that a cache wants to send can eventually be sent
(either to its parent or to its children)
b) Under the assumption that a parent will eventually "decide" to send a
downgrade request to its children (using ParentSendReq rule) whenever it sees
an upgrade request from a child (as opposed to ignoring such an upgrade request
from a child forever), every request sent to either the parent or to a child
will eventually get a response.

We chose "halting" for pedagogical purposes; in our actual Coq-code we use the
traditional trace-equivalence definition requiring inter-convertible states. It
is worth noting that any safety property can be reduced to the problem of
halting.

We believe such a modular technique for verifying full-system correctness with
respect to SC will be useful for hardware designers/verification-engineers to
independently verify different components of the system. We can't be sure that
everything will generalize effortlessly to the trickiest relaxed memory models
(like those exhibited by ARM/POWER as pointed by Reviewer-C because of no
verification loads and/or shared store buffers), but it would have been hard to
figure out the basic framework if we started on more complex examples first.
We still think our example is quite complex and interesting, which Reviewer-D
seems to agree with. Moreover, the performance benefits of a relaxed memory
model over SC is being actively debated in the architecture community.

Reviewer-C, please note that verification loads do indeed ensure correctness.
This is true even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. Our final
theorem ensures this; otherwise the speculative system will not be an
implementation of the reference SC.

Reviewer-B, please note that our SC model is exactly as specified by Lamport.
Also please note that by local buffers in Figure 2, we do not mean just store
buffers; it includes all local data accessible to the processor.

### 2014-12-15 Rebuttal.txt

Some of the questions raised by the reviewers (including the justification of
modularization advantages) are addressed by Reviewer-D.

We have submitted the source code of our Coq development as a supplement to our
paper in the submission website.

### 2014-12-15 Rebuttal.txt

Some of the questions raised by the reviewers (including the justification of
modularization advantages) are addressed by Reviewer-D.

We have submitted the source code of our Coq development as a supplement to our
paper in the submission website.

In response to "points against" by Reviews B and C:
We believe such a modular technique for verifying full-system correctness with
respect to SC will be useful for hardware designers/verification-engineers to
independently verify different components of the system. We can't be sure that
everything will generalize effortlessly to the trickiest relaxed memory models
(like those exhibited by ARM/POWER as pointed by Reviewer-C because of no
verification loads and/or shared store buffers), but it would have been hard to
figure out the basic framework if we started on more complex examples first.
We still think our example is quite complex and interesting, which Reviewer-D
seems to agree with, and our focus is on developing a general framework,
not this particular specification or proof.

Reviewer-C, please note that verification loads do indeed ensure SC correctness.
This is true even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. Our final
theorem ensures this; otherwise the speculative system will not be an
implementation of the reference SC.

Reviewer-B, please note that our SC model is exactly as specified by Lamport
(though it may be written in different notation).  Also please note that by local
buffers in Figure 2, we do not mean just store buffers; it includes all local
data accessible to the processor.

>>>> QUESTION FOR MURALI: What in the reviews does the following paragraph respond to?
     Is it just about the Coq proofs?  (If so, apologies for reordering)
     To me it seems perilously close to just restating our paper abstract. <<<<<

@@

We have structured the proofs modularly as described in the paper. Thus, the
proofs of the different components can be reused while changing the other
components. For example, the proof of the cache-coherence protocol can be
reused as is for verifying other (non-sequentially consistent) systems.
Moreover, the proof of the cache-coherence is itself modular and can be adopted
to other protocols (like MESI, MOSI); The proofs for Lemmas 2 to 5 will change,
but the lemmas correspond to at most 3 components (and states) and hence can be
easily proved (even using techniques like model-checking).

We agree with Reviewer-D that proving liveness is important and non-trivial.
The example of ParentSendReq is an example of a voluntary rule, and the policy
to fire such a rule does indeed affect the liveness. In our development, we
haven't proved full-system liveness (which includes the liveness of the
processor), but we have proved the liveness of the cache-coherence protocol, in
particular, the following:
a) Any request (or response) that a cache wants to send can eventually be sent
(either to its parent or to its children)
b) Under the assumption that a parent will eventually "decide" to send a
downgrade request to its children (using ParentSendReq rule) whenever it sees
an upgrade request from a child (as opposed to ignoring such an upgrade request
from a child forever), every request sent to either the parent or to a child
will eventually get a response.

We chose "halting" for pedagogical purposes; in our actual Coq-code we use the
traditional trace-equivalence definition requiring inter-convertible states. It
is worth noting that any safety property can be reduced to the problem of
halting.

Thanks for notifying us of other related work (Hu et.al.'s "Decidable SC"
SPAA'03, Qadeer's "Verifying SC by model-checking" TPDS'03, Sorin's
"Verification-aware Architectures" PACT'07, "Lamport Clocks: Verifying
Directory Cache-Coherence Protocol" SPAA'98 and Moore's "ACL2 proof of cache
coherence" CAV'98), and for pointing out the other bugs in our related work
section.

Reviewer-B's concerns:
- Ideal processor and memory: As described in that paragraph - memory responds
  instantaneously and processor executes atomically without speculation
- Bluespec served as the model: The motivation for LTS comes from using
  Bluespec to describe hardware, which is similar to our LTS formalization.
- Bluespec's competitive performance can be seen from [Dave et al. 2005], and
  other papers using Bluespec designs.
- Peek-beneath abstraction boundaries: To prove the full-system, the
  individual components are treated as black-boxes, completely captured by their
  interface specification.  Peeking beneath boundaries would instead look inside
  implementations.
- Good-ideas from the world of PL semantics: Using LTSes as a precise hardware
  specification.
- Infinite families of hardware design resist reduction to finite-state
  models: reduction is applied to 'infinite families'
- Figure 2: logical view of system as 3-components: It is more-or-less accepted
  in the architecture community.
- Refining vs Equivalent: Throughout the paper, we are showing only refinement
  in one direction, though you are correct that in this particular example they
  are equivalent.
- Finer-grained architectural details: As mentioned in that paragraph (like
  register-renaming).
- Main-memory still exposes atomic loads and stores: Most architectures (including
  POWER and ARM) can be modeled using a memory that responds to loads and stores
  instantaneously, though the response (or request) can sit in a buffer between
  the memory and the processor, leading to relaxed memory models.

### 2014-12-15 Rebuttal.txt

Some of the questions raised by the reviewers (including the justification of
modularization advantages) are addressed by Reviewer-D.

We have submitted the source code of our Coq development as a supplement to our
paper in the submission website.

In response to "points against" by Reviews B and C:
We believe such a modular technique for verifying full-system correctness with
respect to SC will be useful for hardware designers/verification-engineers to
independently verify different components of the system. We can't be sure that
everything will generalize effortlessly to the trickiest relaxed memory models
(like those exhibited by ARM/POWER as pointed by Reviewer-C because of no
verification loads and/or shared store buffers), but it would have been hard to
figure out the basic framework if we started on more complex examples first.
We still think our example is quite complex and interesting, which Reviewer-D
seems to agree with.

@@

Reviewer-C, please note that verification loads do indeed ensure SC correctness.
This is true even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. Our final
theorem ensures this; otherwise the speculative system will not be an
implementation of the reference SC.

Regarding Reviewer-D's question about reusability of our proofs and framework,
the proof of the cache-coherence protocol can be reused as is for verifying
other (relaxed memory) systems. Moreover, the proof of the cache-coherence is
itself modular and can be adopted to other protocols (like MESI, MOSI); The
proofs for Lemmas 2 to 5 will change, but the lemmas correspond to at most 3
components (and states) and hence can be easily proved (even using techniques
like model-checking).

### 2014-12-15 Rebuttal.txt

Some of the questions raised by the reviewers (including the justification of
modularization advantages) are addressed by Reviewer-D.

We have submitted the source code of our Coq development as a supplement to our
paper in the submission website.

In response to "points against" by Reviews B and C:
We believe such a modular technique for verifying full-system correctness with
respect to SC will be useful for hardware designers/verification-engineers to
independently verify different components of the system. We can't be sure that
everything will generalize effortlessly to the trickiest relaxed memory models
(like those exhibited by ARM/POWER as pointed by Reviewer-C because of no
verification loads and/or shared store buffers), but it would have been hard to
figure out the basic framework if we started on more complex examples first.
We still think our example is quite complex and interesting, which Reviewer-D
seems to agree with, and our focus is on developing a general framework,
not this particular specification or proof.

@@

Reviewer-C, please note that verification loads do indeed ensure SC correctness.
This is true even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. Our final
theorem ensures this; otherwise the speculative system will not be an
implementation of the reference SC.

We chose "halting" for pedagogical purposes; in our actual Coq-code we use the
traditional trace-equivalence definition requiring inter-convertible states. It
is worth noting that any safety property can be reduced to the halting problem.

Thanks for notifying us of other related work (Hu's "Decidable SC" SPAA'03,
Qadeer's "Verifying SC by model-checking" TPDS'03, Sorin's "Verification-aware
Architectures" PACT'07, "Lamport Clocks: Verifying Directory Protocol" SPAA'98
and Moore's "ACL2 cache coherence proof" CAV'98), and for pointing out other
bugs in our related work section.

Beyond 500 words:
Reviewer-B's concerns:
- Ideal processor and memory: As described in that paragraph - memory responds
  instantaneously and processor executes atomically without speculation
- Bluespec served as the model: The motivation for LTS comes from using
  Bluespec to describe hardware, which is similar to our LTS formalization.
- Bluespec's competitive performance can be seen from [Dave et al. 2005], and
  other papers using Bluespec designs.
- Peek-beneath abstraction boundaries: To prove the full-system, the
  individual components are treated as black-boxes, completely captured by their
  interface specification.  Peeking beneath boundaries would instead look inside
  implementations.
- Good-ideas from the world of PL semantics: Using LTSes as a precise hardware
  specification.
- Infinite families of hardware design resist reduction to finite-state
  models: reduction is applied to 'infinite families'
- Figure 2: logical view of system as 3-components: It is more-or-less accepted
  in the architecture community.
- Refining vs Equivalent: Throughout the paper, we are showing only refinement
  in one direction, though you are correct that in this particular example they
  are equivalent.
- Finer-grained architectural details: As mentioned in that paragraph (like
  register-renaming).
- Main-memory still exposes atomic loads and stores: Most architectures (including
  POWER and ARM) can be modeled using a memory that responds to loads and stores
  instantaneously, though the response (or request) can sit in a buffer between
  the memory and the processor, leading to relaxed memory models.

### 2015-01-20 Rebuttal2.txt

1. Why waste time describing LTSes:
We introduced LTSes mainly to define the notion of refinement that we
are using, as well as to define the combination operators.

2. The user of a theorem-prover has to walk the "protocol space" through case analysis:
The theorem-prover has to walk the protocol space through case analysis, but
the proof-script automation in Coq significantly reduces the burden. One can
write proof-scripts that performs a proof-search, i.e. tries to apply
previously proven basic lemmas on each of the cases and leaves the user with
very few actual proof obligations that are complex or unique so as to not be
discharged using automated proof-search.

3. Invariants are the hardest part in parametric verification of a cache
coherence protocol, and we do not mention it:
We do mention the high level theorems that we use in the paper.  While we
restrict ourselves to the verification of the MSI protocol's implementation,
the proof is modular, and a large portion of the proof can be used, as is, for
verification of other protocols (like MESI, MOSI and MOESI), as we discuss in
the previous rebuttal. The theorem prover helps construct and prove the high
level theorems that make these protocols correct. These theorems depend on a
few basic properties pertaining that can be established easily, but were not
mentioned in the paper due to lack of space. These properties include a) state
changes occuring when and only when response messages are sent, b) absence of
response messages from the parent to a child and from the same child to the
parent simultaneously in the network, for the same address and c) FIFO
properties of the messages. These basic properties seem to be the foundation
over which the other cache coherence protocols lie -- they eliminate the corner
cases which make the proof difficult. For example, since responses from a child
to its parent and from the parent to the same child for the same address cannot
be present simultaneously, an upgrade response from the parent cannot cross a
downgrade response from the child, which would have eventually led to the
upgrade of the child and downgrade of the parent violating Lemma 3. As pointed
out by Reviewer-E, the theorems and these properties serve as a method to
derive a correct protocol (the different protocols are different system
instances which obey these properties).

4. Pipelining protocol engines and Bluespec:
Though the semantics of the atomic transitions in our paper execute one at a
time, different independent transitions can be made to occur in parallel in the
actual implementation without changing the semantics; Bluespec already does
this optimization. One can take this further, where each atomic transition can
be spread across multiple hardware clock cycles, while ensuring that the
overall atomic transaction semantics are obeyed. Bluespec currently does not
implement this, though there has been work in this area [Karczmarek et.al.].

5. Nirav's point: I don't see that Reviewer-E's scathing concern fall
into the core of what we're addressing; our focus on modularity is to
allow verification to be (mostly) carried over to different
microarchitectural variants, which is a bare minimum if one wants
architects to have any lateral freedom to design a working
implementation with a low-cost path to verification. I don't see why
there's any fundamental reason that "scalable" approaches he's pointed to
can't be fit into our methodology directly? I've only glanced at the
papers, but it seems like it should be so.
--- Note sure how to condense it and incorporate -Murali

Reviewer-F:
1. Evidence of proofs being machine-checked:
We have submitted the source code of our Coq development as a supplement to our
paper in the submission website.

2. Figure 3, Tags in Figures 5 & 6:
Both these points are valid and we could indicate these points explicitly.

3. Figure 6: The third case:
Reviewer-F's assertion is not true. The "upd" function merely determines if the
"real" state of the system is as dictated by the state (s,pc) or if it should
be modified to take into account the responses from the memory system. If the
processor is not waiting for the memory system response, then the state is
exactly (s,pc). We made an optimization in the fourth case where we make the
real state H whenever we know that the next instruction we decode is H; instead
the state can be left as (s,pc) in the fourth case.

4. Figure 8: Tags for non-speculative loads:
Our invariant is that there can be only one outstanding non-speculative load,
though there can be several outstanding speculative loads along with the
outstanding non-speculative load. Thus, anything without a tag is the
non-speculative load and there is only one of them, and anything with a tag is
a speculative load.

Reviewer-E's related work summary:

FM methodology: The methodology used for proving that the memory system
realizes cache coherence is not new. A theorem-proving approach to
verifying memory systems has been tacit in many publications. Researchers
preferred model checking, as opposed to pushing through theorem-proving
based proofs largely because of the daunting complexity of realistic
memory system protocols. One can find several approaches that address
specific aspects of memory system verification in, say, Wim Hesselink's
work : http://www.informatik.uni-trier.de/~ley/pers/hd/h/Hesselink:Wim_H=.

@@@ Murali response
Lots of PVS-based proofs for concurrent algorithms, mutual exclusion, etc...
(Needs to be read further, but it's just a bunch of mechanized proofs for
a lot of concurrent algorithms)
***

There has of course been recent trends to simplify protocols thru the DeNoVo
project:
http://scholar.google.com/citations?view_op=view_citation&hl=en&user=ufGYoL0AAAAJ&citation_for_view=ufGYoL0AAAAJ:8k81kl-MbHgC
The authors have not cited such related pieces of work and justified the
simplicity of their design.

@@@ Murali response
A new non-conventional cache coherence protocol that produces sequentially
consistent results for data-rate free programs. There's no formal proof that
this protocol is correct...
(Needs to be read further, but doesn't seem relevant)
***

The authors claim that they are achieving their proof for an arbitrary number
of processors. They do not mention anything about the invariants involved
in their parametric proofs. This leads me to suspect that these invariants
are pretty straightforward.  If one harks back to work addressing real
protocols, the Chou, Mannava, Park work
(http://scholar.google.com/citations?view_op=view_citation&hl=en&user=ufGYoL0AAAAJ&citation_for_view=ufGYoL0AAAAJ:qjMakFHDy7sC) shows how exceedingly hard
it is to discover these invariants. In fact, the CMP approach is an
incremental CEGAR-style method for discovering these invariants. Stating these
invariants explicitly would be virtually impossible - even for FLASH-like
protocols.  In fact, a whole series of papers have tried to address invariant
discovery in this space : http://dl.acm.org/citation.cfm?id=1517434

@@@ Murali response
Chou, Mannava, Park [CMP-method] approach uses three paradigms: a) symmetry of
cache coherence protocols b) abstraction of several nodes which behaves as just
another node and c) counter-example guided abtraction refinement (CEGAR) for
refining the abstraction's properties. Using CEGAR, they get invariants for the
rest of the system while concentrating on a finite number of nodes (2).  Each
refinement will not restrict the behavior, but will help towards proving a
safety property. Specifically, let P be an invariant for the abstract
transition system, and we should prove that it satisfies safety property X. We
start out with a P that doesn't satisfy X. Then we add constraints on P, while
permitting the same transitions. Eventually P will satisfy X. The manual
portion of this procedure is in strengthening P.
***

@@@ Murali response
The second paper tries to formulate better invariants in the first place using
the notion of "flows". It says what other messages can interfere when the other
message is being processed...
(Needs to be read further -- I don't understand "flows" clearly; can't say
anything more before I undestand that properly)
***

For hierarchical protocols, this work has been extended in e.g.,
http://link.springer.com/article/10.1007%2Fs10703-010-0092-y

@@@ Murali response
Same approach as [CMP-method] to hierarchical protocols.
***

### 2015-01-20 RelatedWork.tex
stuff

Cache-coherence proofs have also been done with mechanized theorem provers. The
correctness of the FLASH cache-coherence protocol~\cite{flash} for a
single-level hierarchy was proven with PVS~\cite{park} using a mapping, called
``aggregation of transactions,'' from multiple localized transactions to a
single global transaction, and finally using model checking to prove the
correctness of the global transactions model. However, they do not provide a
methodology to prove the aggregation of transactions for other protocols, which
is where, we believe, the difficulty lies; in contrast, our modular proof
relies on generic invariants, making it more adaptable to different protocols.
Cachet~\cite{StoyShenArvind:Proofs}, another single-level cache coherence
protocol, was verified to be obey the CRF memory model~\cite{Shen:CRF}, and
parts of the proof were implemented using PVS.

>>>>

Cache-coherence proofs have also been done with mechanized theorem provers. The
correctness of the FLASH cache-coherence protocol~\cite{flash} for a
single-level hierarchy was proven with PVS~\cite{park} using a mapping, called
``aggregation of transactions,'' from multiple localized transactions to a
single global transaction, and finally using model checking to prove the
correctness of the global transactions model. However, they do not provide a
methodology to prove the aggregation of transactions for other protocols, which
is where, we believe, the difficulty lies; in contrast, our modular proof
relies on generic invariants, making it more adaptable to different protocols.
Cachet~\cite{StoyShenArvind:Proofs}, another single-level cache coherence
protocol, was verified to obey the CRF memory model~\cite{Shen:CRF}, and
parts of the proof were implemented using PVS.
